---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)

## Packages to use ----
pacman::p_load(
  tidyverse, janitor, writexl,
  readxl, scales, mytidyfunctions
)

## Set theme ------
theme_set(theme_jmr(text = element_text(family = "Times New Roman")))

options(
  ggplot2.discrete.colour = c("#277DA1", "#f94144", "#F9C74F", "#43AA8B"),
  ggplot2.discrete.fill = c("#277DA1", "#f94144", "#F9C74F", "#43AA8B")
)
```

# rbrm: Regularized Binary Regression Models for Relative Risk

<!-- badges: start -->
<!-- badges: end -->

## Overview

`rbrm` provides a comprehensive framework for estimating **relative risks** (RR) using penalized binary regression models. Unlike traditional logistic regression (which estimates odds ratios), this package directly models relative risk through a novel log odds-product parameterization, avoiding the limitations of Poisson regression that can produce invalid probabilities.

## Installation

You can install the development version from [GitHub](https://github.com/JavierMtzRdz/rbrm):

``` r
# Install remotes if needed
# install.packages("remotes")
remotes::install_github("JavierMtzRdz/rbrm")
```

## Quick Start

```{r example, eval=T}
library(rbrm)

# Simulate data
set.seed(123)
n <- 200
p <- 5

data <- generate_data(
  n = n,
  pa = p,
  pb = p,
  n_test = 100
)

# Fit regularized model (Penalized)
fit <- rbrm(
  va = data$v.train,
  vb = data$v.train,
  x = data$x.train,
  y = data$y.train,
  lambda = 0.02,
  adjusted = FALSE
)

fit

# Cross-validation with custom metric and adjusted final model
cv_fit <- cv_rbrm(
  data$v.train,
  vb = data$v.train,
  x = data$x.train,
  y = data$y.train,
  measure = "deviance" # Options: deviance, auc, brier, misclass
)

cv_fit
plot(cv_fit)
cv_fit$final_fit

# Fit relaxed regularization path (no refitting)
fit_path <- rbrm(
  va = data$v.train,
  vb = data$v.train,
  x = data$x.train,
  y = data$y.train,
  adjusted = FALSE
)

fit_path
plot(fit_path) +
  geom_vline(xintercept = cv_fit$lambda_min,
             alpha = 0.3, linetype = "dashed")

# Fit relaxed regularization path (Unpenalized Refitting)
fit_path <- rbrm(
  va = data$v.train,
  vb = data$v.train,
  x = data$x.train,
  y = data$y.train,
  adjusted = TRUE
)

fit_path
plot(fit_path) 
```

## The Model

This package implements the Richardson-Robins-Wang (2017) binary regression model for relative risk:

- **Outcome Model**: $P(Y=1 \mid X, W, Z) = p_1$ if $X=1$, $p_0$ if $X=0$
- **RR Parameterization**: $\log(\text{RR}) = \theta = W^\top \alpha$ (treatment effect)
- **Baseline Risk**: $p_0$ via log odds-product $\phi = Z^\top \beta$

The model directly estimates relative risk while ensuring valid probabilities $(0 \leq p_0, p_1 \leq 1)$.

### Regularization

Supports L1 (lasso) and elastic net penalties for high-dimensional data:

- Variable selection when $p > n$
- Separate penalties for $\alpha$ and $\beta$ parameters
- Unpenalized intercept option

## Available Optimizers

1. **FISTA** (default): Fast proximal gradient with momentum and adaptive restart
2. **Newton-CD**: Coordinate descent with Newton steps and diagonal Hessian
3. **Active Set Newton-CD**: Efficient for sparse solutions via KKT conditions
4. **L-BFGS**: Quasi-Newton method (best for low-dimensional, unpenalized problems)

## References

Richardson, T. S., Robins, J. M., & Wang, L. (2017). On modeling and estimation for the relative risk and risk difference. *Journal of the American Statistical Association*, 112(519), 1121-1130. [arXiv:1510.02430](https://arxiv.org/abs/1510.02430)

## Acknowledgments

This package extends the original [regularized-RR-regression](https://github.com/ChloeYou/regularized-RR-regression) implementation and builds upon the [`brm`](https://github.com/mclements/brm) package.

## License

GPL (>= 2)
